---
title: "Lab 03 - Nobel laureates"
author: "Fiona Wang"
date: "2025-02-03"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

There are `r nrow(nobel)` rows in the dataset, meaning there are `r nrow(nobel)` observations in total. There are `r ncol(nobel)` variables in the dataset. Each row gives the information of one Nobel laureates, specifically, their names and country of birth, death, and prize-winning. 


### Exercise 2

```{r filter_dataset}
nobel_living = nobel %>% 
  filter(!is.na(country) & gender != "org" & is.na(died_date))
```

This new data frame has `r nrow(nobel_living)` rows.   

Next, we create a new variable: country_us.

```{r country_us}
nobel_living <- nobel_living %>% 
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

We will limit our analysis to only Physics, Medicine, Chemistry, and Economics. 
```{r limit_categories}
nobel_living_science <- nobel_living %>% 
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

It's interesting that we still have `r nrow(nobel_living_science)` observations. 

### Exercise 3
```{r bar_plot}
nobel_living_science %>% 
  ggplot(mapping = aes(y = country_us, color=category, fill = category)) + 
  geom_bar() + facet_wrap(~category, nrow = 2) + 
  labs(title = "Born in the U.S vs. Others", subtitle = "by category")
```

According to the visualization, I do observe that more Nobel laureate were based in the U.S. when they won the Nobel prize for all four categories. This is more true for Economics than other categories. I would conclude that the visualization supports Buzzfeed headline, but I can't be sure whether the difference between the bars is significant. 

### Exercise 4
```{r born_country_us}
nobel_living_science <- nobel_living_science %>% 
  mutate(
    born_country_us = if_else(born_country == "USA", "USA", "Other")
  )
nobel_living_science %>% 
  count(born_country_us)
```

Of the 228 winners, there are 105 that were born in the U.S.

### Exercise 5
```{r born_base_us}
nobel_living_science %>% 
  ggplot(mapping = aes(x = country_us, fill = born_country_us)) + 
  geom_bar() + facet_wrap(~category, ncol = 2) + 
  labs(title = "Based vs. Born", subtitle = "by category")
```

Of those US-based Nobel winners, there are actually more US born winners. We are looking at the bars that say USA (on the right of each facet). There are more area of blue (born in USA), so this data do NOT support Buzzfeed's claim, acknolwedging that we cannot conduct significance testing.

### Exercise 6

...
