---
title: "Lab 03 - Nobel laureates"
author: "Fiona Wang"
date: "2025-02-03"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

There are `r nrow(nobel)` rows in the dataset, meaning there are `r nrow(nobel)` observations in total. There are `r ncol(nobel)` variables in the dataset. Each row gives the information of one Nobel laureates, specifically, their names and country of birth, death, and prize-winning. 


### Exercise 2

```{r filter_dataset}
nobel_living = nobel %>% 
  filter(!is.na(country) & gender != "org" & is.na(died_date))
```

This new data frame has `r nrow(nobel_living)` rows.   

### Exercise 3

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...
