---
title: "Lab 03 - Nobel laureates"
author: "Fiona Wang"
date: "2025-02-03"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

There are `r nrow(nobel)` rows in the dataset, meaning there are `r nrow(nobel)` observations in total. There are `r ncol(nobel)` variables in the dataset. Each row gives the information of one Nobel laureates, specifically, their names and country of birth, death, and prize-winning. 


### Exercise 2

```{r filter_dataset}
nobel_living = nobel %>% 
  filter(!is.na(country) & gender != "org" & is.na(died_date))
```

This new data frame has `r nrow(nobel_living)` rows.   

Next, we create a new variable: country_us.

```{r country_us}
nobel_living <- nobel_living %>% 
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

We will limit our analysis to only Physics, Medicine, Chemistry, and Economics. 
```{r limit_categories}
nobel_living_science <- nobel_living %>% 
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

It's interesting that we still have `r nrow(nobel_living_science)` observations. 

### Exercise 3
```{r bar_plot}
nobel_living_science %>% 
  ggplot(mapping = aes(y = country_us, color=category, fill = category)) + 
  geom_bar() + facet_wrap(~category, nrow = 2) + 
  labs(title = "Born in the U.S vs. Others", subtitle = "by category")
```

According to the visualization, I do observe that more Nobel laureate were based in the U.S. when they won the Nobel prize for all four categories. This is more true for Economics than other categories. I would conclude that the visualization supports Buzzfeed headline, but I can't be sure whether the difference between the bars is significant. 

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...
